### 2017-12-10 resources/text_as_data/tc4.tex

\documentclass{beamer}

%\usepackage[table]{xcolor}
\mode<presentation> {
  \usetheme{Boadilla}
%  \usetheme{Pittsburgh}
%\usefonttheme[2]{sans}
\renewcommand{\familydefault}{cmss}
%\usepackage{lmodern}
%\usepackage[T1]{fontenc}
%\usepackage{palatino}
%\usepackage{cmbright}
  \setbeamercovered{transparent}
\useinnertheme{rectangles}
}
%\usepackage{normalem}{ulem}
%\usepackage{colortbl, textcomp}
\setbeamercolor{normal text}{fg=black}
\setbeamercolor{structure}{fg= black}
\definecolor{trial}{cmyk}{1,0,0, 0}
\definecolor{trial2}{cmyk}{0.00,0,1, 0}
\definecolor{darkgreen}{rgb}{0,.4, 0.1}
\usepackage{array}
\newcommand{\argmin}{\arg\!\min}
\beamertemplatesolidbackgroundcolor{white}  \setbeamercolor{alerted
text}{fg=red}

\setbeamertemplate{caption}[numbered]\newcounter{mylastframe}

%\usepackage{color}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usepackage{colortbl}
%\usepackage[usenames, dvipsnames]{color}
%\setbeamertemplate{caption}[numbered]\newcounter{mylastframe}c
%\newcolumntype{Y}{\columncolor[cmyk]{0, 0, 1, 0}\raggedright}
%\newcolumntype{C}{\columncolor[cmyk]{1, 0, 0, 0}\raggedright}
%\newcolumntype{G}{\columncolor[rgb]{0, 1, 0}\raggedright}
%\newcolumntype{R}{\columncolor[rgb]{1, 0, 0}\raggedright}

%\begin{beamerboxesrounded}[upper=uppercol,lower=lowercol,shadow=true]{Block}
%$A = B$.
%\end{beamerboxesrounded}}
\renewcommand{\familydefault}{cmss}
%\usepackage[all]{xy}

\usepackage{tikz}
\usepackage{lipsum}

 \newenvironment{changemargin}[3]{%
 \begin{list}{}{%
 \setlength{\topsep}{0pt}%
 \setlength{\leftmargin}{#1}%
 \setlength{\rightmargin}{#2}%
 \setlength{\topmargin}{#3}%
 \setlength{\listparindent}{\parindent}%
 \setlength{\itemindent}{\parindent}%
 \setlength{\parsep}{\parskip}%
 }%
\item[]}{\end{list}}
\usetikzlibrary{arrows}
%\usepackage{palatino}
%\usepackage{eulervm}
\usecolortheme{lily}
\newtheorem{com}{Comment}
\newtheorem{lem} {Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{obs}{Observation}
 \numberwithin{equation}{section}

%\usepackage[latin1]{inputenc}
\title[Text as Data] % (optional, nur bei langen Titeln nötig)
{Text as Data}

\author{Justin Grimmer}
\institute[Stanford University]{Associate Professor\\Department of Political Science \\  Stanford University}
\vspace{0.3in}

\date{October 2nd, 2014}%[Big Data Workshop] 
%\date{\today}

\begin{document}
\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\frametitle{Classification via Dictionary Methods}

\begin{itemize}
\item[1)] Task 
\begin{itemize} 
\invisible<1>{\item[a)] Categorize documents into predetermined categories}
\invisible<1-2>{\item[b)] Measure documents association with predetermined categories} 
\end{itemize}
\invisible<1-3>{\item[2)] Objective function:} 
\begin{eqnarray}
\invisible<1-4>{f(\boldsymbol{\theta}, \boldsymbol{X}_{i}) & = & \frac{\sum_{j=1}^{N} \alert<7>{\theta}_{j} \alert<8>{X}_{ij} }{\sum_{j = 1}^{N} \alert<8>{X}_{ij} } \nonumber } 
\end{eqnarray}

\invisible<1-5>{where: } 
\begin{itemize}
\invisible<1-6>{\item[-] $\boldsymbol{\theta} = (\theta_{1}, \theta_{2}, \hdots, \theta_{N})$ are word weights}
\invisible<1-7>{\item[-] $\boldsymbol{X}_{i} = (X_{i1}, X_{i2}, \hdots, X_{iN})$ count the occurrence of each corresponding word in document $i$}
\end{itemize}

\invisible<1-8>{\item[3)] Optimization$\leadsto$ predetermined word list, no task specific optimization}

\invisible<1-9>{\item[4)] Validation (Model checking)$\leadsto$ weight (model) checking, replication of hand coding, face validity}

\end{itemize}

\pause \pause \pause \pause \pause \pause \pause \pause \pause

\end{frame}

\begin{frame}
\frametitle{Word Weights: Separating Classes}

General Classification Goal: Place documents into categories\pause \\

\invisible<1>{How To Do Classification?} \pause

\begin{itemize}
\invisible<1-2>{\item[-] Dictionaries:} \pause 
\begin{itemize}
\invisible<1-3>{\item[-] Rely on Humans$\leadsto$ humans to identify words that associate with classes} \pause 
\invisible<1-4>{\item[-] Measure how well words separate (positive/negative, emotional, ...)} \pause 
\end{itemize}
\invisible<1-5>{\item[-] Supervised Classification Methods (Later in the Quarter): } \pause 
\begin{itemize}
\invisible<1-6>{\item[-] Rely on statistical models} \pause 
\invisible<1-7>{\item[-] Given set of coded documents, statistical relationship between classes/words} \pause 
\invisible<1-8>{\item[-] Statistical measures of separation} \pause 
\end{itemize}
\end{itemize}

\large 
\invisible<1-9>{Key point: this is the same task}

\end{frame}

\begin{frame}
\frametitle{Types of Classification Problems}

\alert{Topic}: What is this text about? \pause 
\invisible<1>{\begin{itemize}
\item[-] Policy area of legislation  \\
$\Rightarrow$ $\{$Agriculture, Crime, Environment, ...$\}$
\item[-] Campaign agendas \\ 
$\Rightarrow$ $\{$Abortion, Campaign, Finance, Taxing, ...       $\}$
\end{itemize}} \pause

\invisible<1-2>{\alert{Sentiment}: What is said in this text? [\alert{Public Opinion}] } \pause 
\invisible<1-3>{\begin{itemize}
\item[-] Positions on legislation\\
 $\Rightarrow$ $\{$ Support, Ambiguous, Oppose $\}$
\item[-] Positions on Court Cases \\
$\Rightarrow$ $\{$ Agree with Court, Disagree with Court $\}$
\item[-] Liberal/Conservative Blog Posts \\
$\Rightarrow$ $\{$ Liberal, Middle, Conservative, No Ideology Expressed $\}$
\end{itemize} } \pause

\invisible<1-4>{\alert{Style}/\alert{Tone}: How is it said?} \pause 
\invisible<1-5>{\begin{itemize}
\item[-] Taunting in floor statements\\
 $\Rightarrow$ $\{$ Partisan Taunt, Intra party taunt, Agency taunt, ... $\}$
\item[-] Negative campaigning \\
$\Rightarrow$ $\{$ Negative ad, Positive ad$\}$
\end{itemize} }

\end{frame}

\begin{frame}
\frametitle{Pre-existing word weights$\leadsto$ Dictionaries}

\invisible<1>{{\tt DICTION}}\\

\invisible<1>{\only<2>{\scalebox{0.55}{\includegraphics{DICTION2.png}}}} 
\only<3>{\scalebox{0.55}{\includegraphics{DICTION3.png}}} 
\only<4>{\scalebox{0.55}{\includegraphics{DICTION4.png}}} 
\only<5>{\scalebox{0.85}{\includegraphics{DICTION5.png}}} 
\only<6>{\scalebox{0.85}{\includegraphics{DictionCost.png}}}

\pause \pause \pause \pause \pause

\end{frame}

\begin{frame}

\scalebox{0.75}{\includegraphics{Year.jpg}}

\end{frame}

\begin{frame}
\frametitle{Dictionary Methods}

Many Dictionary Methods (like DICTION) \pause

\begin{itemize}
\invisible<1>{\item[1)] Proprietary}\pause\invisible<1-2>{$\leadsto$ wrapped in GUI} \pause 
\invisible<1-3>{\item[2)] Basic tasks:} \pause
\begin{itemize}
\invisible<1-4>{\item[a)] Count words} \pause 
\invisible<1-5>{\item[b)] Weighted counts of words} \pause 
\invisible<1-6>{\item[c)] Some graphics}\pause 
\end{itemize}
\invisible<1-7>{\item[3)] Pricey$\leadsto$ \alert{inexplicably}} 
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{DICTION}

\begin{columns}[]

\column{0.5\textwidth}
\scalebox{0.15}{\includegraphics{PolTone.jpg}}

\column{0.5\textwidth}
\pause 
\begin{itemize}
\item[-] \invisible<1>{$\{$ Certain, Uncertain $\}$}\pause\invisible<1-2>{\\, $\{$ Optimistic, Pessimistic $\}$} \pause
\item[-] \invisible<1-3>{$\approx$ 10,000 words} \pause 
\end{itemize}

\invisible<1-4>{Applies DICTION to a wide array of political texts\\} \pause 
\invisible<1-5>{Examine specific periods of American political history}

\end{columns}

\end{frame}

\begin{frame}
\frametitle{Other Dictionaries }

\begin{itemize}
\item[1)] General Inquirer Database (\url{http://www.wjh.harvard.edu/~inquirer/} ) \pause 
\begin{itemize}
\invisible<1>{\item[-] Stone, P.J., Dumphy, D.C., and Ogilvie, D.M. (1966) \emph{The General Inquirer: A Computer Approach to Content Analysis}} \pause 
\invisible<1-2>{\item[-] $\{$ Positive, Negative $\}$ } \pause 
\invisible<1-3>{\item[-] 3627 negative and positive word strings } \pause 
\invisible<1-4>{\item[-] Workhorse for classification across many domains/papers} \pause 
\end{itemize}
\invisible<1-5>{\item[2)] Linguistic Inquiry Word Count (LIWC)} \pause 
\begin{itemize}
\invisible<1-6>{\item[-] Creation process:} \pause 
\begin{itemize}
\invisible<1-7>{\item[1)] Generate word list for categories$\leadsto$ `` We drew on common emotion rating scales...Roget's Thesaurus...standard English dictionaries. [then] brain-storming sessions among 3-6 judges were held" to generate other words } \pause 
\invisible<1-8>{\item[2)] Judge round$\leadsto$ (a) Does the word belong? (b) What other categories might it belong to?} \pause 
\end{itemize}
\invisible<1-9>{\item[-] $\{$ Positive emotion, Negative emotion $\}$} \pause 
\invisible<1-10>{\item[-] 2300 words grouped into 70 classes} \pause 
\end{itemize}
\invisible<1-11>{\item[-] Harvard-IV-4 } \pause 
\invisible<1-12>{\item[-] Affective Norms for English Words (we'll discuss this more later)} \pause 
\invisible<1-13>{\item[-] ...} 
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Generating New Words}

Three ways to create dictionaries (non-exhaustive): \pause 
\begin{itemize}
\invisible<1>{\item[-] Statistical methods$\leadsto$ next Tuesday} \pause 
\invisible<1-2>{\item[-] Manual generation } \pause 
\begin{itemize}
\invisible<1-3>{\item[-] Careful thought (prayer? epiphanies? divine intervention?) about useful words} \pause 
\end{itemize}
\invisible<1-4>{\item[-] Populations of people who are surprisingly willing to perform ill-defined tasks} \pause 
\begin{itemize}
\invisible<1-5>{\item[a)] Undergraduates$:\text{Pizza}\rightarrow \text{Research Output}$} \pause 
\invisible<1-6>{\item[b)] Mechanical turkers} \pause 
\begin{itemize}
\invisible<1-7>{\item[-] Example: $\{$ Happy, Unhappy $\}$ } \pause 
\invisible<1-8>{\item[-] Ask turkers: how happy is } \pause 
\invisible<1-9>{\item[] {\tt elevator}, {\tt car}, {\tt pretty}, {\tt young} } \pause 
\invisible<1-10>{\item[] Output as dictionary} 
\end{itemize}
\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Applying Methods to Documents}

Applying the model: \pause 
\begin{itemize}
\invisible<1>{\item[-] Vector of word counts:  $\boldsymbol{X}_i = (X_{i1}, X_{i2}, \hdots, X_{iK}$, $(i = 1, \hdots, N)$} \pause
\invisible<1-2>{\item[-] Weights attached to words  $\boldsymbol{\theta} = (\theta_{1}, \theta_{2}, \hdots, \theta_{K})$  } \pause 
\begin{itemize}
\invisible<1-3>{\item[-] $\theta_{k} \in \{0,1\}$} \pause
\invisible<1-4>{\item[-] $\theta_{k} \in \{-1, 0, 1 \}$} \pause
\invisible<1-5>{\item[-] $\theta_{k} \in \{-2, -1, 0, 1, 2\}$} \pause
\invisible<1-6>{\item[-] $\theta_{k} \in \Re$} \pause
\end{itemize}
\end{itemize}

\invisible<1-7>{For each document $i$ calculate score for document } \pause
\begin{eqnarray}
\invisible<1-8>{Y_i  & = &  \frac{\sum_{k=1}^{K} \theta_k X_{ik}}{\sum_{k=1}^{K} X_{k}} \nonumber \\} \pause
\invisible<1-9>{Y_i  & = &  \frac{\boldsymbol{\theta}^{'} \boldsymbol{X}_i}{\boldsymbol{X}_{i}^{'} \boldsymbol{1} } \nonumber } \pause
\end{eqnarray}

\invisible<1-10>{$Y_{i} \approx $ continuous $\leadsto$ Classification} \pause
\begin{itemize}
\invisible<1-11>{\item[] $Y_i> 0 \Rightarrow$ Positive Category} \pause
\invisible<1-12>{\item[] $Y_i< 0 \Rightarrow$ Negative Category} \pause
\invisible<1-13>{\item[] $Y_i \approx 0$ Ambiguous} 
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Applying a Dictionary to Press Releases}

\pause 
\begin{itemize}
\invisible<1>{\item[-] Collection of 169,779 press releases (US House members 2005-2010)} \pause 
\invisible<1-2>{\item[-] Dictionary from Neal Caren's website $\leadsto$ Theresa Wilson, Janyce Wiebe, and Paul Hoffman's dictionary } \pause 
\invisible<1-3>{\item[-] Create positive/negative score for press releases.  } 
\end{itemize}

\invisible<1-4>{{\tt Python} code and press releases}

\pause 
\end{frame}

\begin{frame}
\frametitle{Examining Positive and Negative Statements in Press Releases}

\pause

\only<1-10>{
\invisible<1>{Least positive members of Congress:} 
\begin{itemize}
\invisible<1-2>{\item[1)] Dan Burton, 2008} 
\invisible<1-3>{\item[2)] Nancy Pelosi, 2007} 
\invisible<1-4>{\item[3)] Mike Pence 2007} 
\invisible<1-5>{\item[4)] John Boehner, 2009}  
\invisible<1-6>{\item[5)] Jeff Flake, (basically all years)} 
\invisible<1-7>{\item[6)] Eric Cantor, 2009} 
\invisible<1-8>{\item[7)] Tom Price, 2010} 
\end{itemize}

\invisible<1-9>{Legislators who are more extreme$\leadsto$ less positive in press releases}

}

\only<11>{\scalebox{0.5}{\includegraphics{pressOverTime.pdf}}}

\only<12-13>{
\begin{itemize}
\item[-] Credit Claiming press release: 9.1 percentage points ``more positive" than a non-credit claiming press release
\invisible<1-12>{\item[-] Anti-spending press release: 10.6 percentage points ``less positive" than a non-anti spending press release}
\end{itemize}
}

\only<14>{\scalebox{0.5}{\includegraphics{CreditPositive.pdf}}}
\only<15->{\scalebox{0.5}{\includegraphics{AntiCreditPositive.pdf}}}

\pause \pause \pause \pause \pause\pause \pause \pause \pause \pause \pause \pause \pause \pause

\end{frame}

\begin{frame}
\frametitle{Methodological Issues/Problems with Dictionaries}

\alert{Dictionary methods are context invariant} \pause \\
\begin{itemize}
\invisible<1>{\item[-] No optimization step $\leadsto$ same word weights regardless of texts} \pause 
\invisible<1-2>{\item[-] Optimization$\leadsto$ incorporate information specific to context} \pause 
\invisible<1-3>{\item[-] Without optimization$\leadsto$ unclear about dictionaries performance} \pause 
\end{itemize}

\invisible<1-4>{\alert{Just because dictionaries provide measures labeled ``positive" or ``negative" it doesn't mean they are accurate measures in your text} (!!!!) \\} \pause

\vspace{0.5in}

\invisible<1-5>{{\huge \alert{Validation}}}

\end{frame}

\begin{frame}
\frametitle{Validation}

Classification Validity: \pause 
\begin{itemize}
\invisible<1>{\item[-] \alert{Training}: build dictionary on subset of documents \alert{with known labels}} \pause 
\invisible<1-2>{\item[-] \alert{Test}: apply dictionary method to other documents \alert{with known labels}} \pause 
\invisible<1-3>{\item[-] Requires hand coded documents} \pause 
\invisible<1-4>{\item[-] Hand coded documents useful for other reasons} \pause 
\begin{itemize}
\invisible<1-5>{\item[-] Is the classification scheme well defined for your texts?} \pause 
\invisible<1-6>{\item[-] Can humans accomplish the coding task?} \pause 
\invisible<1-7>{\item[-] Is the dictionary your using appropriate?} \pause 
\end{itemize}
\end{itemize}

\large 
\invisible<1-8>{\alert{Replicate} classification exercise}  \pause 
\normalsize 
\begin{itemize}
\invisible<1-9>{\item[-] How well does our method perform on \alert{held out} documents?} \pause 
\invisible<1-10>{\item[-] Why held out?} \pause \invisible<1-11>{\alert{Over fitting} } \pause 
\invisible<1-12>{\item[-] Using off-the-shelf dictionary: all labeled documents to test} \pause 
\invisible<1-13>{\item[-] Supervised learning classification: \alert{(Cross)validation} } 
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Hand Coding: A Brief Digression}

\alert{Humans should be able to classify documents into the categories you want the machine to classify them in} \pause 
\begin{itemize}
\invisible<1>{\item[-] This is \alert{hard}} \pause 
\invisible<1-2>{\item[-] Why? } \pause 
\begin{itemize}
\invisible<1-3>{\item[-] Ambiguity in language} \pause 
\invisible<1-4>{\item[-] Limited working memory} \pause 
\invisible<1-5>{\item[-] Ambiguity in classification rules} \pause 
\end{itemize}
\invisible<1-6>{\item[-] A procedure for training coders: } \pause 
\invisible<1-7>{\begin{itemize}
\item[1)] Coding rules
\item[2)] Apply to new texts
\item[3)] Assess coder agreement (we'll discuss more in a few weeks)
\item[4)] Using information and discussion, revise coding rules
\end{itemize}} 
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Assessing Classification}

Measures of classification performance

\begin{tabular}{l|l|l}
 \hline
  & \multicolumn{2}{c}{Actual Label}  \\
  \hline
  Guess &   Liberal & Conservative \\
  \hline 
  Liberal &  \alert{True Liberal} & False Liberal \\
  \hline
  Conservative & False Conservative & \alert{True Conservative} \\
  \hline
  \hline
\end{tabular}

\pause 
\begin{eqnarray}
\invisible<1>{\text{Accuracy} & = & \frac{ \alert{\text{TrueLib} }+ \alert{\text{TrueCons}}  } { \alert{\text{TrueLib} } + \alert{\text{TrueCons}} + \text{FalseLib} + \text{FalseCons} } \nonumber } \pause  \\
\invisible<1-2>{\text{Precision}_{\text{Liberal}} &= &   \frac{ \alert{\text{True Liberal}}    }  { \alert{\text{True Liberal }} + \text{False Liberal}      } } \pause  \nonumber \\
\invisible<1-3>{\text{Recall}_{\text{Liberal} } & = & \frac{ \alert{\text{True Liberal}}   } { \alert{\text{True Liberal}} + \text{False Conservative}   } } \pause  \nonumber \\
\invisible<1-4>{F_{\text{Liberal}} & = & \frac{ 2\text{Precision}_{\text{Liberal}} \text{Recall}_{\text{Liberal} } } { \text{Precision}_{\text{Liberal}} +  \text{Recall}_{\text{Liberal} }} }   \nonumber \pause 
\end{eqnarray}

\invisible<1-5>{\alert{Under reported for dictionary classification} } 
\end{frame}

\begin{frame}
\frametitle{What about continuous measures?}

\pause

\invisible<1>{\alert{Necessarily more complicated}\\} \pause

\begin{itemize}
\invisible<1-2>{\item[-] Go back to hand coding exercise} \pause 
\invisible<1-3>{\item[-] Imagine asking undergraduates to rate document on a continuous scale (0-100)} \pause 
\invisible<1-4>{\item[-] \alert{Difficult} to create classifications with agreement} \pause 
\invisible<1-5>{\item[-] \alert{Precisely} the point$\leadsto$ merely creating a gold standard is hard, let alone computer classification} \pause 
\end{itemize}

\invisible<1-6>{\alert{Lower level classification}}\pause\invisible<1-7>{$\leadsto$ label phrases and then aggregate} \pause \\

\invisible<1-8>{Modifiable areal unit problem in texts}\pause$\leadsto$\invisible<1-9>{aggregating destroys information, conclusion may depend on level of aggregation}

\end{frame}

\begin{frame}
\frametitle{Validation, Dictionaries from other Fields}
\pause 
\invisible<1>{Accounting Research: measure \alert{tone} of \alert{10-K} reports} \pause 
\begin{itemize}
%\item[-] Comprehensive public summary of company performance
\invisible<1-2>{\item[-] \alert{tone} matters (\$)} \pause 
\end{itemize}

\invisible<1-3>{Previous state of art: Harvard-IV-4 Dictionary applied to texts} \\
\invisible<1-4>{Loughran and McDonald (2011): \alert{Financial Documents are Different}, \textcolor{blue}{polysemes} } \pause 
\begin{itemize}
\invisible<1-5>{\item[-] Negative words in Harvard, Not Negative in Accounting: \\} \pause 
\invisible<1-6>{{\tt tax, cost, capital, board, liability, foreign,  cancer, crude (oil), tire } } \pause 
\invisible<1-7>{\item[-] \alert{73\%} of Harvard negative words in this set(!!!!!)} \pause 
\invisible<1-8>{\item[-] Not Negative Harvard, Negative in Accounting: \\} \pause 
\invisible<1-9>{{\tt felony, litigation, restated, misstatement, and unanticipated} } \pause 
\end{itemize}

\large 
\invisible<1-10>{\alert{Context Matters}}

\end{frame}

\begin{frame}
\frametitle{Measuring Happiness}

\begin{columns}[]
\column{0.5\textwidth}
\scalebox{0.35}{\includegraphics{Bentham.jpg}}

\column{0.5\textwidth}

\pause 
\begin{itemize}
\invisible<1>{\item[-] Quantifying Happiness: How happy is society?} \pause 
\invisible<1-2>{\item[-] How Happy is a Song?} \pause 
\invisible<1-3>{\item[-] Blog posts?} \pause
\invisible<1-4>{\item[-] Facebook posts? (Gross National Happiness)} \pause 
\end{itemize}

\invisible<1-5>{Use \alert{Dictionary Methods} }

\end{columns}

\end{frame}

\begin{frame}
\frametitle{Measuring Happiness}

Dodds and Danforth (2009): Use a dictionary method to measure happiness \pause 
\begin{itemize}
\invisible<1>{\item[-]  \alert{Affective Norms for English Words} (ANEW)} \pause 
\invisible<1-2>{\item[-] Bradley and Lang 1999:  1034 words, Affective reaction to words} \pause 
\begin{itemize}
\invisible<1-3>{\item[-] On a scale of 1-9 how happy does this word make you?} \pause 
\invisible<1-4>{\item[] \alert{Happy} : triumphant (8.82)/paradise (8.72)/ love (8.72) } \pause 
\invisible<1-5>{\item[] \alert{Neutral}: street (5.22)/ paper (5.20)/ engine (5.20) } \pause 
\invisible<1-6>{\item[] \alert{Unhappy} : cancer (1.5)/funeral (1.39)/ rape (1.25) /suicide (1.25) } \pause 
\end{itemize}
\invisible<1-7>{\item[-] \alert{Happiness} for text $i$ (with word $j$ having happiness $\theta_j$ and document frequence $X_{ij}$} \pause 
\begin{eqnarray}
\invisible<1-8>{\text{Happiness}_{i}  & = & \frac{ \sum_{k=1}^{K} \theta_{k} X_{ik} } { \sum_{k=1}^{K} X_{ik}} }  \nonumber 
\end{eqnarray}
\end{itemize}

\end{frame}

\begin{frame}

\scalebox{0.5}{\includegraphics{BillyJean.png}}
\pause

\invisible<1>{\alert{Homework Hints}:}
\invisible<1>{One approach: write a {\tt for} loop searching for words in dictionary (caution: is dictionary stemmed?) }\\ \pause 
\invisible<1-2>{Happiest Song on Thriller?}  \\ \pause 
\invisible<1-3>{\alert{P.Y.T. (Pretty Young Thing) }   (This is the right answer!)}

\end{frame}

\begin{frame}
\frametitle{Happiness in Society}

\only<1>{\scalebox{1}{\includegraphics{SongHappiness.png}}}
\only<2>{\scalebox{1}{\includegraphics{SongType.png}}}
\only<3>{\scalebox{0.7}{\includegraphics{Blog.png}}}

\end{frame}

\begin{frame}
\frametitle{Dictionary Methods}

Today: Classification via Dictionaries\\
Next week: Seperating Words and the Geometry of Text\\
Good luck on the homework!

\end{frame}

\end{document}