### 2017-12-10 resources/text_as_data/tc5.tex

\documentclass{beamer}

%\usepackage[table]{xcolor}
\mode<presentation> {
  \usetheme{Boadilla}
%  \usetheme{Pittsburgh}
%\usefonttheme[2]{sans}
\renewcommand{\familydefault}{cmss}
%\usepackage{lmodern}
%\usepackage[T1]{fontenc}
%\usepackage{palatino}
%\usepackage{cmbright}
  \setbeamercovered{transparent}
\useinnertheme{rectangles}
}
%\usepackage{normalem}{ulem}
%\usepackage{colortbl, textcomp}
\setbeamercolor{normal text}{fg=black}
\setbeamercolor{structure}{fg= black}
\definecolor{trial}{cmyk}{1,0,0, 0}
\definecolor{trial2}{cmyk}{0.00,0,1, 0}
\definecolor{darkgreen}{rgb}{0,.4, 0.1}
\usepackage{array}
\newcommand{\argmin}{\arg\!\min}
\beamertemplatesolidbackgroundcolor{white}  \setbeamercolor{alerted
text}{fg=red}

\setbeamertemplate{caption}[numbered]\newcounter{mylastframe}

%\usepackage{color}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usepackage{colortbl}
%\usepackage[usenames, dvipsnames]{color}
%\setbeamertemplate{caption}[numbered]\newcounter{mylastframe}c
%\newcolumntype{Y}{\columncolor[cmyk]{0, 0, 1, 0}\raggedright}
%\newcolumntype{C}{\columncolor[cmyk]{1, 0, 0, 0}\raggedright}
%\newcolumntype{G}{\columncolor[rgb]{0, 1, 0}\raggedright}
%\newcolumntype{R}{\columncolor[rgb]{1, 0, 0}\raggedright}

%\begin{beamerboxesrounded}[upper=uppercol,lower=lowercol,shadow=true]{Block}
%$A = B$.
%\end{beamerboxesrounded}}
\renewcommand{\familydefault}{cmss}
%\usepackage[all]{xy}

\usepackage{tikz}
\usepackage{lipsum}

 \newenvironment{changemargin}[3]{%
 \begin{list}{}{%
 \setlength{\topsep}{0pt}%
 \setlength{\leftmargin}{#1}%
 \setlength{\rightmargin}{#2}%
 \setlength{\topmargin}{#3}%
 \setlength{\listparindent}{\parindent}%
 \setlength{\itemindent}{\parindent}%
 \setlength{\parsep}{\parskip}%
 }%
\item[]}{\end{list}}
\usetikzlibrary{arrows}
%\usepackage{palatino}
%\usepackage{eulervm}
\usecolortheme{lily}
\newtheorem{com}{Comment}
\newtheorem{lem} {Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{obs}{Observation}
 \numberwithin{equation}{section}

%\usepackage[latin1]{inputenc}
\title[Text as Data] % (optional, nur bei langen Titeln nötig)
{Text as Data}

\author{Justin Grimmer}
\institute[Stanford University]{Associate Professor\\Department of Political Science \\  Stanford University}
\vspace{0.3in}

\date{October 7th, 2014}%[Big Data Workshop] 
%\date{\today}

\begin{document}
\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\frametitle{Estimating Word Discrimination}

\begin{itemize}
\item[1)] Task
\begin{itemize}
\item[a)] \alert{Classification}$\leadsto$ learn word weights for dictionaries
\item[b)] \alert{Fictitious prediction problem}$\leadsto$ Identify features that discriminate between groups to learn features that are indicative of some group
\end{itemize}
\item[2)] Objective function
\begin{eqnarray}
f(\boldsymbol{\theta}, \boldsymbol{X}) &= & f(\boldsymbol{\theta}, \boldsymbol{X}, \boldsymbol{Y}) \nonumber 
\end{eqnarray}
where: 
\begin{itemize}
\item[] $\boldsymbol{Y}$ = Document Labels 
\item[] $\boldsymbol{X}$  = Document Features
\item[] $\boldsymbol{\theta}$ = Parameters that measure words discrimination between categories
\end{itemize}
\item[3)] Optimization$\leadsto$ method specific
\item[4)] Validation $\leadsto$ depends on task
\begin{itemize}
\item[i)] Classification$\leadsto$ Accuracy, Precision, Recall
\item[ii)] Fictitious prediction$\leadsto$ Face, convergent, discriminatory, and \alert{confound}
\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Stylometry$\leadsto$ Who Wrote Disputed Federalist Papers?}

Federalist papers $\leadsto$ Mosteller and Wallace (1963)
\begin{itemize}
\item[-] Persuade citizens of New York State to adopt constitution
\item[-] Canonical texts in study of American politics
\item[-] 77 essays 
\begin{itemize}
\item[-] Published from 1787-1788 in Newspapers
\item[-] And under the name \alert{Publius}, anonymously
\end{itemize}
\end{itemize}
\alert{Who Wrote the Federalist papers}?
\begin{itemize}
\item[-] Jay wrote essays 2, 3, 4,5, and 64
\item[-] Hamilton: wrote 43 papers
\item[-] Madison: wrote 12 papers
\end{itemize}
\alert{Disputed:} Hamilton or Madison?
\begin{itemize}
\item[-] Essays: 49-58, 62, and 63
\item[-] Joint Essays: 18-20
\end{itemize}
\alert{Task}: identify authors of the disputed papers.  \\
\alert{Task}: Classify papers as Hamilton or Madison using dictionary methods

\end{frame}

\begin{frame}
\frametitle{Setting up the Analysis}

\alert{Training}$\leadsto$ papers Hamilton, Madison are known to have authored\\
\alert{Test}$\leadsto$ unlabeled papers\\
\alert{Preprocessing}:
\begin{itemize}
\item[-] Hamilton/Madison both discuss similar issues
\item[-] Differ in extent they use \alert{stop words}
\item[-] Focus analysis on the stop words
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Setting up the Analysis}

\begin{itemize}
\item[-] $\boldsymbol{Y} = (Y_{1}, Y_{2}, \hdots, Y_{N}) = (\text{Hamilton}, \text{Hamilton}, \text{Madison}, ..., \text{Hamilton})  $ \\ $N \times 1$ matrix with author labels
\item[-] Define the number of words in federalist paper $i$ as num$_{i}$
$\boldsymbol{X} = \begin{pmatrix}
\frac{1}{\text{num}_{1}}  & \frac{2}{\text{num}_{1}} & \frac{0}{\text{num}_{1}} & \hdots & \frac{3}{\text{num}_{1}}\\
\frac{0}{\text{num}_{2}} & \frac{1}{\text{num}_{2}} & \frac{0}{\text{num}_{2}} & \hdots & \frac{0}{\text{num}_{2}} \\
\vdots & \vdots & \vdots& \ddots & \vdots \\
\frac{0}{\text{num}_{N}} & \frac{0}{\text{num}_{N}} & \frac{1}{\text{num}_{N}} & \hdots & \frac{0}{\text{num}_{N}} \\
\end{pmatrix}$

 $N \times J$ counting stop word usage rate
\item[-] $\boldsymbol{\theta} = (\theta_{1}, \theta_{2}, \hdots, \theta_{J})$ \\ Word weights.\\
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Objective Function}

\alert{Heuristically}: find $\boldsymbol{\theta}^{*} = (\theta_{1}^{*}, \theta_{2}^{*}, \hdots, \theta_{J}^{*}) $ used to create score 
\begin{eqnarray}
p_{i}  = \sum_{j=1}^{J} \theta_{j}^{*} X_{ij} \nonumber 
\end{eqnarray}

 that maximally discriminates between categories

\scalebox{0.4}{\includegraphics{LDALine.pdf}}

\end{frame}

\begin{frame}
\frametitle{Objective Function}

Define:
\begin{eqnarray}
\boldsymbol{\mu}_{\text{Madison}} & = & \frac{1}{N_{\text{Madison}}} \sum_{i=1}^{N} I(Y_{i} = \text{Madison}) \boldsymbol{X}_{i} \nonumber \\
\boldsymbol{\mu}_{\text{Hamilton}} & = & \frac{1}{N_{\text{Hamilton}}} \sum_{i=1}^{N} I(Y_{i} = \text{Hamilton}) \boldsymbol{X}_{i} \nonumber 
\end{eqnarray}

\end{frame}

\begin{frame}
\frametitle{Objective Function}

\begin{small}

We can then define functions that describe the ``projected" mean and variance for each author

\begin{eqnarray}
g(\boldsymbol{\theta}, \boldsymbol{X}, \boldsymbol{Y},  \text{Madison}) & = & \frac{1}{N_{\text{Madison}}} \sum_{i= 1}^{N}  I(Y_{i} = \text{Madison} ) \boldsymbol{\theta}^{'}\boldsymbol{X}_{i}  = \boldsymbol{\theta}^{'} \boldsymbol{\mu}_{\text{Madison}} \nonumber  \\
g(\boldsymbol{\theta}, \boldsymbol{X}, \boldsymbol{Y},  \text{Hamilton}) & = & \frac{1}{N_{\text{Hamilton}}} \sum_{i= 1}^{N}  I(Y_{i} = \text{Hamilton} ) \boldsymbol{\theta}^{'}\boldsymbol{X}_{i}  = \boldsymbol{\theta}^{'} \boldsymbol{\mu}_{\text{Hamilton}} \nonumber  \\
s(\boldsymbol{\theta}, \boldsymbol{X}, \boldsymbol{Y},  \text{Madison}) & = & \sum_{i=1}^{N} I(Y_{i} = \text{Madison} ) (\boldsymbol{\theta}^{'} \boldsymbol{X}_{i} - \boldsymbol{\theta}^{'} \boldsymbol{\mu}_{\text{Madison}})^2 \nonumber \\
s(\boldsymbol{\theta}, \boldsymbol{X}, \boldsymbol{Y},  \text{Hamilton}) & = & \sum_{i=1}^{N} I(Y_{i} = \text{Hamilton} ) (\boldsymbol{\theta}^{'} \boldsymbol{X}_{i} - \boldsymbol{\theta}^{'} \boldsymbol{\mu}_{\text{Hamilton}})^2 \nonumber 
\end{eqnarray}

\end{small}

\end{frame}

\begin{frame}
\frametitle{Objective Function$\leadsto$ Optimization}

\begin{eqnarray}
f(\boldsymbol{\theta}, \boldsymbol{X}, \boldsymbol{Y} ) & = & \frac{\left(g(\boldsymbol{\theta}, \boldsymbol{X}, \boldsymbol{Y},  \text{Hamilton}) - g(\boldsymbol{\theta}, \boldsymbol{X}, \boldsymbol{Y},  \text{Madison}) \right)^2}{s(\boldsymbol{\theta}, \boldsymbol{X}, \boldsymbol{Y},  \text{Hamilton})  + s(\boldsymbol{\theta}, \boldsymbol{X}, \boldsymbol{Y},  \text{Madison}) } \nonumber\\
& =&  \frac{\left(\boldsymbol{\theta}^{'} ( \boldsymbol{\mu}_{\text{Hamilton}} - \boldsymbol{\mu}_{\text{Madison}})     \right)^2  }{\text{Scatter}_{\text{Hamilton}} + \text{Scatter}_{\text{Madison}}    } \nonumber 
\end{eqnarray}

\alert{Optimization}$\leadsto$ find $\boldsymbol{\theta}^{*}$ to maximize $f(\boldsymbol{\theta}, \boldsymbol{X}, \boldsymbol{Y} ) $, assuming independence across dimensions.\\

\alert{(Fisher's) Linear Discriminant Analysis}

\end{frame}

\begin{frame}
\frametitle{Optimization$\leadsto$ Word Weights}

For each word $j$, construct weight $\theta^{*}_{j}$, 
\begin{small}
\begin{eqnarray}
\mu_{j, \text{Hamilton}} & = & \frac{\sum_{i = 1}^{N} I(Y_{i} = \text{Hamilton}) X_{ij} }{\sum_{j=1}^{J} \sum_{i = 1}^{N} I(Y_{i} = \text{Hamilton}) X_{ij} } \nonumber \\
\mu_{j, \text{Madison}} & = & \frac{\sum_{i = 1}^{N} I(Y_{i} = \text{Madison}) X_{ij} }{\sum_{j=1}^{J} \sum_{i = 1}^{N} I(Y_{i} = \text{Madison}) X_{ij} } \nonumber \\
\sigma^{2}_{j, \text{Hamilton}} & = & \text{Var}(X_{i,j} |  \text{Hamilton}) \nonumber \\
\sigma^{2}_{j, \text{Madison} }& = & \text{Var}(X_{i,j} | \text{Madison}) \nonumber 
\end{eqnarray}
\end{small}

We can then generate weight $\theta^{*}_{j}$ as

\begin{eqnarray}
\theta^{*}_j &= & \frac{\mu_{j, \text{Hamilton}}  - \mu_{j, \text{Madison}}  } {\sigma_{j, \text{Hamilton}}^{2} + \sigma_{j, \text{Madison}}^{2}    }  \nonumber 
\end{eqnarray}

\end{frame}

\begin{frame}
\frametitle{Optimization$\leadsto$ Trimming the Dictionary}

\begin{itemize}
\item[-] Trimming weights: Focus on discriminating words (very simple \alert{regularization})
\item[-] Cut off: For all $|\theta^{*}_{j}|<0.025$ set $\theta^{*}_{j} = 0$.  
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Classification$\leadsto$ Determining Authorship}

For each disputed document $i$, compute discrimination statistic
\begin{eqnarray}
p_{i}  & = & \sum_{j =1}^{J} \theta^{*}_{j} X_{ij} \nonumber 
\end{eqnarray}

$p_i \leadsto$ classification (\alert{linear discriminator})
\begin{itemize}
\item[-] Above midpoint in training set $\rightarrow$ Hamilton text
\item[-] Below midpoint in training set $\rightarrow$ Madison text
\end{itemize}

\alert{Findings}: Madison is the author of the disputed federalist papers.

\end{frame}

\begin{frame}
\frametitle{Inferring Separating Words}

\only<1-9, 11->{Classification$\leadsto$ Custom Dictionaries 
\begin{itemize}
\invisible<1>{\item[-] Stylometry$\leadsto$ Classify Authors} 
\invisible<1-2>{\item[-] Dictionary based classification $\leadsto$ Gentzkow and Shapiro (2010) and measures of media slant} \pause 
\invisible<1-3>{\item[-] Dictionary based classification$\leadsto$ Customized to particular setting}
\end{itemize}
\invisible<1-4>{\alert{Fictitious Prediction Problem} $\leadsto$ Infer words that are indicative of some class/group} 
\begin{itemize}
\invisible<1-5>{\item[-] Difference in Republican, Democratic language $\leadsto$ \alert{Partisan} words} 
\invisible<1-6>{\item[-] Difference in Liberal, Conservative language $\leadsto$ Ideological Language} 
\invisible<1-7>{\item[-] Difference in Secret/Not Secret Language $\leadsto$ Secretive Language (Gill and Spirling 2014)}  
\invisible<1-8>{\item[-] Difference in Toy advertising } 
\invisible<1-10>{\item[-] Difference in Language across groups$\leadsto$ Labeling output from Clustering/Topic Models}
\end{itemize}
\invisible<1-11>{\alert{Vague} and \alert{Difficult} to derive before hand}

}
\begin{center}
\only<10>{\scalebox{0.4}{\includegraphics{BoyGirlsAd.jpg}}}
\end{center}
\pause \pause\pause \pause \pause  \pause \pause \pause \pause \pause

\end{frame}

\begin{frame}
\frametitle{Congressional Language Across Sources}
\pause

\invisible<1>{Congressional Press Releases and Floor Speeches} \pause 
\begin{itemize}
\invisible<1-2>{\item[-] Collected 64,033 press releases} \pause 
\invisible<1-3>{\item[-] Problem: are they \alert{distinct} from floor statements (approx. 52,000 during same time)?} \pause 
\begin{itemize}
\invisible<1-4>{\item[-] Yes: press releases have different purposes, targets, and need not relate to official business } \pause 
\invisible<1-5>{\item[-] No: press releases are just reactive to floor activity, will follow floor statements} \pause 
\end{itemize}
\invisible<1-6>{\item[-] Deeper question: what does it mean for two text collections to be \alert{different?}} \pause 
\invisible<1-7>{\item[-] One Answer: \alert{texts used for different purposes}} \pause 
\invisible<1-8>{\item[-] Partial answer: identify words that distinguish press releases and floor speeches} 
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{A Method for Identifying Distinguishing Words}

\pause

\invisible<1>{Mutual Information} \pause 
\begin{itemize}
\invisible<1-2>{\item[-] Unconditional uncertainty (entropy): } \pause 
\begin{itemize}
\invisible<1-3>{\item[-] Randomly sample a press release } \pause 
\invisible<1-4>{\item[-] Guess press release/floor statement} \pause 
\invisible<1-5>{\item[-] Uncertainty about guess } \pause 
\begin{itemize}
\invisible<1-6>{\item[-] Maximum: No. press releases = No. floor statements} \pause 
\invisible<1-7>{\item[-] Minimum : All documents in one category} \pause 
\end{itemize}
\end{itemize}
\invisible<1-8>{\item[-] Conditional uncertainty ($X_{j}$) (conditional entropy)} \pause 
\begin{itemize}
\invisible<1-9>{\item[-] Condition on presence of word $X_{j}$}\pause
\invisible<1-10>{\item[-] Randomly sample a press release}\pause
\invisible<1-11>{\item[-] Guess press release/floor statement}\pause
\invisible<1-12>{\item[-] Word presence reduces uncertainty }\pause
\begin{itemize}
\invisible<1-13>{\item[-] Unrelated: Conditional uncertainty = uncertainty}\pause
\invisible<1-14>{\item[-] Perfect predictor: Conditional uncertainty = 0}\pause
\end{itemize}
\end{itemize}
\invisible<1-15>{\item[-] Mutual information($X_{j}$): uncertainty - conditional uncertainty ($X_{j}$)} \pause
\begin{itemize}
\invisible<1-16>{\item[-] Maximum: Uncertainty $\rightarrow$ $X_{j}$ is perfect predictor} \pause
\invisible<1-17>{\item[-] Minimum: 0 $\rightarrow$ $X_{j}$ fails to separate speeches and floor statements}
\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{A Method for Identifying Distinguishing Words}

\pause 
\begin{itemize}
\invisible<1>{\item[-] $\text{Pr}(\text{Press}) \equiv $ Probability selected document press release} \pause 
\invisible<1-2>{\item[-] $\text{Pr}(\text{Speech}) \equiv $ Probability selected document speech} \pause 
\invisible<1-3>{\item[-] Define \alert{entropy} $H(\text{Doc})$ } \pause 
\begin{eqnarray}
\invisible<1-4>{H(\text{Doc})  & = &  \alert{-} \sum_{t \in \{\text{Pre}, \text{Spe} \} } \text{Pr}(t) \log_2 \text{Pr}(t) \nonumber } \pause 
% & = &  - (\text{Pr(Press)}\log_2 \text{Pr(Press)} + \text{Pr(Speech)} \log_2 \text{Pr(Speech)} ) \nonumber  
\end{eqnarray}
\invisible<1-5>{\item[-] $\log_2$? Encodes bits} \pause 
\invisible<1-6>{\item[-] Maximum: Pr(Press) = Pr(Speech) = 0.5} \pause 
\invisible<1-7>{\item[-] Minimum: Pr(Press) $\rightarrow$ 0 (or Pr(Press) $\rightarrow$ 1)} 
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{A Method for Identifying Distinguishing Words}

\pause
\begin{itemize}
\invisible<1>{\item[-] Consider presence/absence of word $X_j$} \pause 
\invisible<1-2>{\item[-] Define \alert{conditional entropy} $H(\text{Doc}|X_j)$ } \pause 
\end{itemize}
\begin{eqnarray}
\invisible<1-3>{H(\text{Doc}|X_j) & = &  - \sum_{s = 0}^{1} \sum_{t \in \{\text{Pre}, \text{Spe} \} } \text{Pr}(t, X_j = s) \log_2 \text{Pr}(t| X_j = s) \nonumber } \pause
\end{eqnarray}
\begin{itemize}
\invisible<1-4>{\item[-] Maximum: $X_j$ unrelated to Press Releases/Floor Speeches} \pause
\invisible<1-5>{\item[-] Minimum: $X_j$ is a perfect predictor of press release/floor speech} 
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{A Method for Identifying Distinguishing Words}

\pause

\begin{itemize}
\invisible<1>{\item[-] Define \alert{Mutual Information}($X_j$) as } \pause 
\begin{eqnarray}
\invisible<1-2>{\text{Mutual Information}(X_j) & = & H(\text{Doc}) - H(\text{Doc}|X_j) \nonumber } \pause 
\end{eqnarray}
\invisible<1-3>{\item[-] Maximum: entropy $\Rightarrow H(\text{Doc}|X_j) = 0$} \pause 
\invisible<1-4>{\item[-] Minimum:  0  $\Rightarrow H(\text{Doc}|X_j) = H(\text{Doc}) $.  } \pause 
\end{itemize}
\large
\invisible<1-5>{Bigger mutual information $\Rightarrow$ better discrimination } \pause \\
\vspace{0.25in} 
\invisible<1-6>{Objective function and optimization$\leadsto$ estimate probabilities that we then place in mutual information}

\end{frame}

\begin{frame}
\frametitle{A Method for Identifying Distinguishing Words}

Formula for mutual information \\ (based on ML estimates of probabilities) 
\begin{eqnarray}
n_p & = & \text{Number Press Releases} \nonumber \\
n_s & = & \text{Number of Speeches} \nonumber \\
D  & = & n_p + n_s \nonumber \\
n_j & = & \sum_{i=1}^{D} X_{i,j} \hspace{0.25in} \text{ (No. docs $X_j$ appears ) } \nonumber \\
n_{-j} & =& \text{ No. docs $X_j$ does not appear  } \nonumber \\
n_{j,p} & = & \text{No. press and $X_j$} \nonumber \\
n_{j,s} & = & \text{No. speech and $X_j$} \nonumber \\
n_{-j,p} & = & \text{No. press and not $X_j$} \nonumber \\
n_{-j,s} & = & \text{No. speech and not $X_j$} \nonumber 
\end{eqnarray}

\end{frame}

\begin{frame}
\frametitle{A Method for Identifying Distinguishing Words}

Formula for Mutual Information

\begin{eqnarray}
\text{MI}(X_j)  & = & \frac{n_{j,p}}{D} \log_2 \frac{n_{j,p} D}{n_j n_p}
    +\frac{n_{j,s}}{D} \log_2 \frac{n_{j,s} D}{n_j n_{s}}
    \nonumber \\
   & &
    +\frac{n_{-j,p}}{D} \log_2 \frac{n_{-j,p} D}{n_{-j} n_p}
    +\frac{n_{-j, s}}{D} \log_2 \frac{n_{-j,s} D}{n_{-j} n_{s}}.
     \nonumber
\end{eqnarray}

\end{frame}

\begin{frame}
\frametitle{What's Different About Press Releases}

\only<1>{\scalebox{0.225}{\includegraphics{MutInfPlot1.pdf}}}
\only<2>{\scalebox{0.225}{\includegraphics{MutInfPlot2.pdf}}}
\only<3>{\scalebox{0.225}{\includegraphics{MutInfPlot3.pdf}}} 
\begin{center}
\only<4-10>{\scalebox{0.125}{\includegraphics{MutInfPlot3.pdf}}}
\only<11>{\scalebox{0.45}{\includegraphics{AttentionPressSpeech}}}
\end{center}

\alert{What's Different?}

\only<4->{\begin{itemize}
\invisible<1-4>{\item[-] Press Releases: Credit Claiming} 
\invisible<1-5>{\item[-] Floor Speeches: Procedural Words} 
\invisible<1-6>{\item[-] Validate: Manual Classification}  
\invisible<1-7>{\item[-] Sample 500 Press Releases, 500 Floor Speeches} 
\invisible<1-8>{\item[-] Credit Claiming: 36\% Press Releases, 4\% Floor Speeches} 
\invisible<1-9>{\item[-] Procedural: 0\% Press Releases, 44\% Floor Speeches} 
\invisible<1-10>{\item[-] Validate: Topic Classification} 
\end{itemize}
}

\pause \pause \pause \pause \pause \pause \pause\pause \pause\pause 
\end{frame}

\begin{frame}
\frametitle{Fightin' Words$\leadsto$ An Introduction to Regularization}

Monroe, Colaresi, and Quinn (2009)$\leadsto$ what makes a word partisan? \pause \\

\invisible<1>{Argue for using \alert{Log Odds Ratio}, weighted by variance\\} \pause

\invisible<1-2>{Recall: For some event $E$ and $F$} \pause 
\begin{eqnarray}
\invisible<1-3>{P(E) & = & 1 - P(E^{c}) \nonumber \\} \pause 
\invisible<1-4>{\text{Odds}(E) & = & \frac{P(E)}{1- P(E)} \nonumber \\} \pause 
\invisible<1-5>{\text{Odds Ratio}(E, F) & = & \frac{\frac{P(E))}{(1 - P(E))}}{\frac{P(F)}{1-P(F)}}\nonumber \\} \pause 
\invisible<1-6>{\text{Log Odds Ratio}(E, F) & = & \log\left(\frac{P(E)}{1- P(E)}\right) - \log\left(\frac{P(F)}{1-P(F)} \right) \nonumber} \pause 
\end{eqnarray}

\invisible<1-7>{Strategy$\leadsto$ Construct objective function on $\alert{proportions}$ (and then calculate log-odds)} \pause

\end{frame}

\begin{frame}
\frametitle{Objective Function}

Suppose we're interested in how a word separates partisan speech.  \\
$\boldsymbol{Y} = (\text{Republican}, \text{Republican}, \text{Democrat}, \hdots, \text{Republican})$\\
$\boldsymbol{X} =$ Unnormalized matrix of word counts $N \times J$

Define
\begin{small}
\begin{eqnarray}
\boldsymbol{x}_{\text{Republican}} & = & (\sum_{i=1}^{N}I(Y_{i} = \text{Republican})X_{i1}, \sum_{i=1}^{N}I(Y_{i} = \text{Republican})X_{i2}, \nonumber \\
&&  \hdots, \sum_{i=1}^{N}I(Y_{i} = \text{Republican})X_{iJ} ) \nonumber 
\end{eqnarray}
\end{small}

with $N_{\text{Republican}} = $ Total number of Republican words

\end{frame}

\begin{frame}
\frametitle{Objective Function}

\pause 
\begin{eqnarray}
\invisible<1>{\boldsymbol{\pi}_{\text{Republican}} & \sim & \text{Dirichlet}(\boldsymbol{\alpha}) \nonumber \\} \pause 
\invisible<1-2>{\boldsymbol{x}_{\text{Republican}}| \boldsymbol{\pi}_{\text{Republican}}  & \sim & \text{Multinomial}(N_{\text{Republican}}, \boldsymbol{\pi}_{\text{Republican}}) \nonumber }\pause 
\end{eqnarray}

\invisible<1-3>{This implies an objective function on $\boldsymbol{\pi}$, } \pause 
\begin{eqnarray}
\invisible<1-4>{p(\boldsymbol{\pi}| \boldsymbol{\alpha}, \boldsymbol{X}, \boldsymbol{Y}) & \propto & p(\boldsymbol{\pi}| \boldsymbol{\alpha}) p(\boldsymbol{x}_{\text{Republican}}| \boldsymbol{\pi} \boldsymbol{\alpha}, \boldsymbol{Y})  \nonumber \\} \pause 
\invisible<1-5>{& \propto & \frac{\Gamma(\sum_{j=1}^{J} \alpha_{j} ) }{\prod_{j} \Gamma(\alpha_{j}) } \prod_{j=1}^{J} \pi_{j}^{\alpha_{j} - 1} \pi_{j}^{x_{\text{Republican}, j}} \nonumber } \pause 
\end{eqnarray}

\invisible<1-6>{$p(\boldsymbol{\pi}| \boldsymbol{\alpha}, \boldsymbol{X}, \boldsymbol{Y}) $ is a Dirichlet distribution:} \pause 
\begin{eqnarray}
\invisible<1-7>{\pi_{\text{Republican}, j}^{*} & = & \frac{x_{\text{Republican}, j} + \alpha_{j} }{N_\text{Republican} + \sum_{j=1}^{J}\alpha_{j}} \nonumber } \pause 
\end{eqnarray}

\end{frame}

\begin{frame}
\frametitle{Calculating Log Odds Ratio}

Define $\log \text{Odds Ratio}_{j} $ as 
\begin{eqnarray}
\log \text{Odds Ratio}_{j} & = & \log \left(\frac{\pi_{\text{Republican}, j}}{1 - \pi_{\text{Republican}, j}} \right) - \log \left( \frac{\pi_{\text{Democratic}, j}}{1 - \pi_{\text{Democratic}, j}} \right) \nonumber 
\end{eqnarray}

\begin{eqnarray}
\text{Var}(\log \text{Odds Ratio}_j ) & \approx & \frac{1 } { x_{jD} + \alpha_j }  + \frac{1} {x_{jR} + \alpha_j }  \nonumber \\
\text{Std. Log Odds}_j & = & \frac { \log \text{Odds Ratio}_j  } { \sqrt{\text{Var}(\log \text{Odds Ratio}_j ) }} \nonumber 
\end{eqnarray}

\end{frame}

\begin{frame}
\frametitle{Applying the Model}

How do Republicans and Democrats differ in debate?\\

Condition on \alert{topic} and examine word usage\\

\begin{itemize}
\item[-] Press Releases (64,033)
\item[-] Topic Coded (Structural Topic Model)
\item[-] Given press release is about topic, what are the features that distinguish Republican and Democratic language?
\end{itemize}

\end{frame}
\begin{frame}

Mutual Information, Standardized Log Odds
\begin{center}
\only<1>{\scalebox{0.375}{\includegraphics{IraqWarPartisan.pdf}}}
\only<2>{\scalebox{0.375}{\includegraphics{GasPricesPartisan.pdf}}}
\end{center}

\end{frame}

\begin{frame}
\frametitle{Multinomial Inverse Regression}

\begin{itemize}
\item[-] In classification we're generally interested in:
\begin{eqnarray}
E[Y| \boldsymbol{X}] & = & g(X_{1}, X_{2}, \hdots, X_{J}) \nonumber 
\end{eqnarray}
\item[-] Problem: $J$ might be very, very big.  
\item[-] Potential solution$\leadsto$ invert regression
\begin{eqnarray}
E[\boldsymbol{X}|Y] & = & g(Y) \nonumber 
\end{eqnarray}
\item[-] Inversion is particularly useful for \alert{feature selection}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Multinomial Inverse Regression: Objective Function (Taddy 2014)}

As before, $\boldsymbol{x}_{\text{Republican}}$ to be the Republican count vector. \pause

\begin{eqnarray}
\invisible<1>{\boldsymbol{x}_{\text{Republican}} & \sim & \text{Multinomial}(N_{\text{Republican}}, \boldsymbol{\pi}_{\text{Republican}}) \nonumber \\} \pause 
\invisible<1-2>{\pi_{\text{Republican}, j} & = & \frac{\exp[\alpha_{j} + I(\text{Republican}) \phi_{j} ]}{\sum_{l=1}^{J} \exp[\alpha_{l} + I(\text{Republican}) \phi_{l} ]}\nonumber \\} \pause 
\invisible<1-3>{\phi_{j} & \sim & \text{Laplace}(\lambda_{j}) \nonumber \\} \pause 
\invisible<1-4>{\lambda_{j} & \sim & \text{Gamma}(s, r) \nonumber } \pause 
\end{eqnarray}

\invisible<1-5>{\alert{Laplace} priors $\leadsto$ \alert{regularize} or \alert{shrink} estimates toward zero \\} \pause 
\invisible<1-6>{\alert{Laplace} priors $\leadsto$ Equivalent to $L1$ or \alert{lasso} penalization\\} \pause 
\invisible<1-7>{\alert{Gamma-Lasso} prior \\} \pause

\invisible<1-8>{Optimization$\leadsto$ Coordinate descent (paper is great!) $\leadsto$ {\tt textir} package}

\end{frame}

\begin{frame}
\frametitle{Applying Multinomial Inverse Regression: Objective Function}

Taddy (2014) considers speeches made on Congressional floor in 2005\\
``Most" Republican words  \pause \\
\invisible<1>{{\tt un.official,term.care.insurance, weapons.grade.plutonium}\\
{\tt million.illegal.immigrant, \alert{grand ole opry}, ..., personal.injury.lawyer}                \\} \pause

\invisible<1-2>{``Most" Democratic Words\\} \pause 
\invisible<1-3>{{\tt wild.bird, dealth.penalty.system, record.budget.deficit}\\
{\tt security.private.account, able.buy.gun}  }

\end{frame}

\end{document}